---
title: "Causal inference with conjoint data"
subtitle: "Day 3, part 1 --- Exercise"
bibliography: references.json
toc: true
---

## Overview

In this exercise, we'll use data from @TeeleKallaRosenbluth2018 to explore how to use conjoints for causal inference. We'll use frequentist methods here, but later today we'll do this again with more complex hierarchical/multilevel Bayesian models.

### Data

The data is in `data/candidate.rds`, and it's a pre-cleaned version of the replication data from @TeeleKallaRosenbluth2018. The original data [@Kalla2018] is [accessible at Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/FVCGHC). I used `00_clean-data.R` to clean it up for this workshop.

In this experiment, respondents were shown pairs of hypothetical political candidates and asked which they would prefer. The candidate profiles varied randomly along 6 different attributes:

| Attribute / feature | Column              | Levels                                             |
|:--------------------|:--------------------|:---------------------------------------------------|
| Gender              | `gender`            | Male, female                                       |
| Years in politics   | `years_politics`    | 0, 1, 3, 8                                         |
| Spouse occupation   | `spouse_occupation` | Unmarried, doctor, farmer                          |
| Occupation          | `occupation`        | Teacher, corporate lawyer, mayor, state legislator |
| Number of children  | `n_children`        | 0, 1, 3                                            |
| Age                 | `age`               | 29, 45, 65                                         |

Each respondent was offered 3 sets of profiles, and each respondent had a randomly assigned level of office for the profiles: Congress, governor, and city council. Respondents were also asked to report their own sex and party identification.

Here's what one hypothetical pairing might look like:

::: {.callout-note title="Example conjoint survey question" icon=false}

If you had to choose between these two candidates **for Congress**, which would you vote for?

| Attribute          | Candidate 1 | Candidate 2 |
|--------------------|:-----------:|:-----------:|
| Gender             |   Female    |   Female    |
| Years in politics  |      1      |      8      |
| Spouse occupation  |  Unmarried  |   Doctor    |
| Occupation         |    Mayor    |   Teacher   |
| Number of children |      1      |      0      |
| Age                |     29      |     45      |
| Choice             |      ○      |      ○      |

:::

## Get a feel for the data

```{r}
#| label: load-libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(patchwork)
library(scales)
library(marginaleffects)
library(parameters)

options(
  parameters_cimethod = FALSE,
  width = 120
)

# From the Urban Institute bc why not
# https://urbaninstitute.github.io/graphics-styleguide/#color
party_colors <- c("Democrat" = "#1696d2", "Republican" = "#db2b27")

# Little helper function for formatting numbers as percentage points
label_pp <- scales::label_number(
  accuracy = 1,
  scale = 100,
  suffix = " pp.",
  style_negative = "minus"
)

# Some lookup tables to help make all the attributes and levels plot in the right order
attributes_levels <- tribble(
  ~attribute, ~attribute_nice, ~level,
  "gender", "Gender", c("Male", "Female"),
  "years_politics", "Years in politics", c("None", "1 year", "3 years", "8 years"),
  "spouse_occupation", "Spouse occupation", c("Unmarried", "Doctor", "Farmer"),
  "occupation", "Occupation", c("Teacher", "Corporate lawyer", "Mayor", "State legislator"),
  "n_children", "Number of children", c("0", "1", "3"),
  "age", "Age", c("29", "45", "65")
)

attribute_lookup <- attributes_levels |> 
  select(-level) |> 
  mutate(attribute_nice = fct_inorder(attribute_nice))

attribute_levels_lookup <- attributes_levels |>
  unnest(level) |> 
  mutate(level_nice = level) |> 
  mutate(across(c("attribute_nice", "level_nice"), \(x) fct_inorder(x)))

candidates <- readRDS("data/candidate.rds")

glimpse(candidates)
```

::: {.callout-tip title="Your turn 1"}
Use your intuition and the code chunk below to answer these questions:

1. What does one row represent?
2. How many respondents are there? How many tasks per respondent? How many profiles per task?
3. The `choice` column is 0 or 1. What does that mean in a forced-choice design? What should `mean(choice)` be, and why?
4. What is `task`? What is `task_id`? Why do we need `task_id` instead of just `task`?
5. How many possible profiles are there (i.e. combinations of levels across features)?
:::

```{r}
#| label: your-turn-1

# Do stuff here
```

::: {.callout-note title="Your turn 1 answers"}
1. **One row = one candidate profile shown to one respondent in one task.** This is the fundamental unit of long conjoint data: respondent × task × alternative.
2. **2,105 respondents; 3 tasks each × 2 profiles per task = 6 possible rows per respondent, or 2,105 × 6 = 12,630 possible rows.** In reality, there are only 12,450 rows, since a handful of respondents only completed 1 or 2 profiles. 
3. **`choice = 1` means that profile was chosen; `choice = 0` means it was rejected.** Each task has exactly one choice, so exactly half the rows must have `choice = 1`, so `mean(choice)` should be 0.5.
4. **`task` is the within-respondent task number (1, 2, or 3).** The same value "1" appears for every respondent's first task, so it does *not* uniquely identify a choice set across the dataset. `task_id` is a globally unique identifier for each choice event for each respondent (which two profiles competed against each other for this respondent). This is important for conditional logit, which needs to know which rows were in the same forced comparison.
5. **2 genders × 4 possible years in politics × 3 spouse occupations × 4 occupations × 3 possible numbers of children × 3 ages = 2 * 4 * 3 * 4 * 3 * 3 = 864 potential profiles**
:::

| Attribute / feature | Column              | Levels                                             |
|:--------------------|:--------------------|:---------------------------------------------------|
| Gender              | `gender`            | Male, female                                       |
| Years in politics   | `years_politics`    | 0, 1, 3, 8                                         |
| Spouse occupation   | `spouse_occupation` | Unmarried, doctor, farmer                          |
| Occupation          | `occupation`        | Teacher, corporate lawyer, mayor, state legislator |
| Number of children  | `n_children`        | 0, 1, 3                                            |
| Age                 | `age`               | 29, 45, 65                                         |

```{r}
#| label: your-turn-1-answers
#| collapse: true

# How many respondents?
candidates |>
  summarize(n_resp = n_distinct(resp_id))

# Rows per respondent. Some didn't do all the tasks!
candidates |>
  count(resp_id) |>
  count(n, name = "n_respondents")

# Mean of choice
candidates |>
  summarize(avg_choice = mean(choice))

# task is 1-3, not unique across respondents
candidates |> count(task)

# task_id is unique per choice set
candidates |>
  summarize(n_task_ids = n_distinct(task_id))
#> # A tibble: 1 × 1
#>   n_task_ids
#>        <int>
#> 1       6225

# Possible profiles
2 * 4 * 3 * 4 * 3 * 3

# Or more programmatically
candidates |>
  summarize(
    across(
      c(gender, years_politics, spouse_occupation, occupation, n_children, age),
      \(x) nlevels(x)
    )
  ) |> 
  mutate(combinations = reduce(c_across(everything()), `*`))
```

Unlike the candy example from Day 1, this experiment has one additional nested layer, since respondents were shown multiple tasks (`task`). I find that it's helpful to visualize the hierarchy of each respondents' six rows in @fig-candidate-hierarchy. Also unlike the candy example, this is not a full factorial design---it would be impossible to show respondents 864 profiles simultaneously! Instead, this uses a *fractional factorial design*. Respondents see random pairs of the 864 profiles, and through the magic of randomization, we can still estimate causal effects for the different attribute levels since they appear uniformly across respondent tasks.

::: {#fig-candidate-hierarchy}

![](img/candidate-multilevel-structure.svg){width="100%"}

Multilevel experimental structure, with candidate choices $\\{y_1, y_2\\}$ nested in tasks in respondents

:::

## General estimands

In general, there are two types of estimands that you can find using conjoint designs: causal and descriptive [@LeeperHoboltTilley2020]---but *nota bene*: this is different from the much richer marketing-style preference description we'll cover tomorrow! The exact research question you have determines what quantity of interest you'll want to find and how you analyze the data to find it [@LundbergJohnsonStewart2021]. @tbl-estimands summarizes the two types of estimands, how they're calculated, and shows some example questions you can answer with them, and we'll explore both in more depth below.

:::: {.column-page-inset}

::: {#tbl-estimands .list-table aligns="l,c,l,l"}

General types of estimands you can calculate for causal inference-focused conjoint analysis

- - Purpose
  - Estimand
  - Example interpretation
  - Practical calculation

- - **Preference description**

    *(but not as fancy as marketing-style preference description!)*

  - **Marginal mean (MM)**

    $$
    \theta = \textbf{E}[\text{Outcome} \mid \text{Feature = Level of interest}]
    $$

    *or*

    $$
    \begin{aligned}
    \theta =\ &\textbf{E}[\text{Outcome} \mid \text{Feature = Level of interest}]\ - \\
    &\textbf{E}[\text{Outcome} \mid \text{Feature = Other level of interest}]
    \end{aligned}
    $$

  - Candidates who are women are supported by X% of respondents
    
    *or*

    Women candidates are X percentage points less likely to be selected than candidates who are corporate lawyers
  
  - Average outcome conditional on attribute levels

- - **Causal inference**

  - **Average marginal component effect (AMCE)**

    $$
    \theta = \mathbf{E}[Y \mid \operatorname{do}(\text{level of interest})] - \mathbf{E}[Y \mid \operatorname{do}(\text{reference})]
    $$

  - Changing a candidate's age from 29 to 45 increases the probability of selection by X percentage points

  - Regression coefficients, marginal effects relative to a reference level, or differences in marginal means

:::

::::

## Marginal means

The marginal mean for an attribute level is the predicted probability of selecting that level, averaging over all other attributes---or holding all other attributes constant, or marginalizing across all other attributes. 

In conjoint designs, marginal means are descriptive and answer questions like "How often would a candidate with this trait be selected, on average?" Importantly, marginal means are independent of reference categories---there is no baseline category to worry about. Instead we can say things like "Female candidates are selected X% of the time, on average."

We compute marginal means in two steps: (1) fit a regression model, and then (2) use `avg_predictions()` to calculate the average predicted probability for one level while holding everything else constant.

```{r}
#| label: base-model
model_lm <- lm(
  choice ~ gender +
    years_politics +
    spouse_occupation +
    occupation +
    n_children +
    age,
  data = candidates
)
```

```{r}
#| label: mm-gender

# Plug in existing data
avg_predictions(model_lm, by = "gender", vcov = ~resp_id)

# Plug in a balanced grid of covariates; in this case, a data frame of all 
# 864 possible profiles
avg_predictions(model_lm, newdata = "balanced", by = "gender", vcov = ~resp_id)
```

::: {.callout-tip title="Your turn 2"}
1. Compute marginal means for all six candidate attributes. Which attribute shows the largest spread across its levels? Which shows the smallest?
2. Make a coefficient plot for all attributes with facets. Add a dashed vertical line at 0.5 (chance selection).
3. What does it mean if two levels of an attribute have very similar marginal means? What does a large spread imply?
:::

```{r}
#| label: your-turn-2

# Do stuff here
```

::: {.callout-note title="Your turn 2 answers"}
1. **Years in politics, candidate occupation, and spouse occupation show the largest spread; number of children, gender, and age show the smallest.** Candidates with 8 years of experience are noticeably preferred over first-timers; mayors and state legislators are preferred over teachers (and voters *really* don't like lawyers!). Voters are fairly indifferent to the number of children and they treat 29-year-olds and 65-year-olds with the same (more negative) favorability, preferring 45-year-old candidates.
2. See plot code below.
3. **Similar marginal means mean respondents are roughly indifferent between those levels**---the attribute barely moves the needle on candidate appeal. A large spread means the attribute strongly shapes who gets selected.
:::

```{r}
#| label: your-turn-2-answers
#| code-fold: true

# We need to calculate predictions for each attribute in isolation, then combine
# them. We could do it like this, but this requires a bunch of data wrangling
# work and it's gross.
# bind_rows(
#   avg_predictions(model_lm, by = "gender", vcov = ~resp_id) |>
#     mutate(attribute = "Gender"),
#   avg_predictions(model_lm, by = "years_politics", vcov = ~resp_id) |>
#     mutate(attribute = "Years in politics"),
#   ...
# )

# So we use some map() magic from {purrr} instead
mm_all <- attribute_lookup |>
  mutate(
    results = map(
      attribute,
      \(att) {
        avg_predictions(model_lm, by = att, vcov = ~resp_id) |>
          # avg_predictions() names the column with the "by" argument, like
          # "gender", "age", etc., so we need to rename it to something
          # consistent for when the rows all get combined
          rename(level = all_of(att))
      }
    )
  ) |>
  unnest(results) |>
  left_join(
    attribute_levels_lookup,
    by = join_by(attribute, attribute_nice, level)
  )

ggplot(mm_all, aes(x = estimate, y = level_nice)) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey50") +
  scale_x_continuous(labels = label_percent()) +
  labs(
    x = "Marginal mean",
    y = NULL,
    title = "Marginal means for all candidate attributes"
  ) +
  facet_wrap(vars(attribute_nice), scales = "free_y", ncol = 2)
```

## Average marginal component effects (AMCEs)

This distinction between causal and descriptive estimands makes sense if we look at the notation for the estimands themselves. [In the world of do-calculus](https://www.andrewheiss.com/blog/2021/09/07/do-calculus-backdoors/), causal questions are asked using the $\operatorname{do}()$ operator, which represents [a direct intervention into a data generating process](https://stats.stackexchange.com/questions/211008/dox-operator-meaning). In this case, the researcher randomly sets the attributes to specific levels—the respondent does not self-select into different conditions or decide for themselves that Candidate 1 is a 45-year-old teacher with 3 kids or that Candidate 2 is a 29-year-old mayor with no kids. This thus eliminates selection bias and other external confounding, leaving us with an average causal effect. 

We can calculate (1) the average outcome when a feature is set to a level we're interested in (e.g., when gender = Female), (2) the average outcome when a feature is set to some reference level (e.g., when gender = Male), and (3) find the difference between the two:

$$
\begin{aligned}
\theta =\ &P [\text{Candidate selection} \mid \operatorname{do}(\text{Gender} = \text{Female})]\ - \\
&P[\text{Candidate selection} \mid \operatorname{do}(\text{Gender} = \text{Male})]
\end{aligned}
$$

Practically speaking, the easiest way to think about the average marginal component effect (AMCE) is as categorical coefficients in a linear regression model. But that's not always necessary---they can also just be contrasts (or differences) in marginal means between an attribute level and its reference level.

In fact, most of the time in many social science disciplines, people will just use raw regression coefficients from a linear probability model as AMCEs:

```{r}
model_parameters(model_lm)
```

In practice, though, it's better to use {marginaleffects} to work with these AMCEs, since it can handle non-linear terms, interaction terms, and nonlinear link functions like logistic regression, more complex multilevel models (which we'll do later!), *and* it can handle clustered standard errors.

```{r}
# Unlike avg_predictions, we can find multiple independent comparisons 
# simultaneously with variables
avg_comparisons(
  model_lm,
  variables = c("gender", "years_politics", "occupation"),
  vcov = ~resp_id
)
```

Switching a candidate from male to female causes a 4.25 percentage point increase in the probability of selection. Switching from no experience in politics to 1 year causes a 5 percentage point increase, and the effect is bigger with more years of experience: ≈10 percentage points for 3 years, ≈15 percentage points for 8 years.

::: {.callout-tip title="Your turn 3"}
1. Calculate all the AMCEs at once with `avg_comparisons()`. Which attribute has the largest AMCE in absolute terms? Does this match the largest spread you found in the marginal means?
2. Interpret the AMCEs for age, occupation, spouse occupation, and number of children. Are any of these effects surprising?
:::

```{r}
#| label: your-turn-3

# Do stuff here
```

::: {.callout-note title="Your turn 3 answers"}
1. **Years in politics has the largest AMCEs.** Eight years of experience vs. no experience produces a huge positive effect (15 percentage points!). This matches the largest spread in the marginal means.
2. **Age**: Middle-aged candidates (45) are strongly preferred---about 8 percentage points more likely to be selected than 29-year-olds. Sixty-five-year-olds are statistically indistinguishable from 29-year-olds (−1 pp, p = 0.187), so voters penalize both youth and old age, preferring candidates in their prime. **Occupation**: People *hate* lawyers---they're about 7 pp *less likely* to be selected than teachers. Mayors and state legislators each gain about 5 pp over teachers. Voters reward relevant government experience. **Spouse occupation**: Having any spouse boosts favorability over being unmarried---farmers by 8 pp, doctors by 6 pp. **Number of children**: Having 1 or 3 children each add about 5 pp over no children, and the effects are nearly identical---voters reward having children but are indifferent to how many.
:::

```{r}
#| label: your-turn-3-answers
#| code-fold: true

all_amces <- avg_comparisons(
  model_lm,
  variables = c(
    "gender",
    "years_politics",
    "spouse_occupation",
    "occupation",
    "n_children",
    "age"
  ),
  vcov = ~resp_id
)

reference_rows <- all_amces |>
  separate_wider_delim(contrast, delim = " - ", names = c("level", "ref")) |>
  select(term, level = ref) |>
  distinct() |>
  mutate(estimate = 0, conf.low = 0, conf.high = 0)

plot_data <- all_amces |>
  separate_wider_delim(contrast, delim = " - ", names = c("level", "ref")) |>
  bind_rows(reference_rows) |> 
  left_join(attribute_levels_lookup, by = join_by(term == attribute, level))

ggplot(plot_data, aes(x = estimate, y = level_nice)) +
  geom_vline(xintercept = 0) +
  geom_pointrange(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.15)
  ) +
  scale_x_continuous(labels = label_pp) +
  labs(x = "AMCE", y = NULL) +
  facet_wrap(vars(attribute_nice), scales = "free_y", ncol = 2) +
  theme(legend.position = "bottom")
```

## Oh no! 1: Subgroup analysis and conditional effects

When analyzing conjoint experiments, we often want to know whether different groups of respondents weight attributes differently. Do men and women, or Democrats and Republicans, or respondents with more or less education evaluate candidate gender or occupation or experience differently?

It feels intuitive to fit a model with interactions between experiment attributes and respondent characteristics and then calculate conditional AMCEs, or AMCEs for each attribute within the different respondent subgroups. But @LeeperHoboltTilley2020 show that this is misleading! Comparing conditional AMCEs across subgroups confounds two things: (1) genuine differences in preferences for a given attribute level and (2) differences in baseline preferences for the reference category. The solution is to compare **conditional marginal means** (differences in favorability on an absolute scale) rather than conditional AMCEs (differences in favorability relative to a hidden baseline).

For instance, let's compare the effect of candidate gender on the probability of selection among Democratic and Republican respondents. First we make a model that interacts respondent party with all the attributes:

```{r}
model_party <- lm(
  choice ~ (gender + years_politics + spouse_occupation +
            occupation + n_children + age) * resp_party,
  data = candidates
)
```

That model has a ton of coefficients, so we'll use {marginaleffects} to build the AMCEs we care about instead of trying to piece together everything from the raw results:

```{r}
conditional_amces_party <- avg_comparisons(
  model_party,
  variables = "gender",
  by = "resp_party",
  vcov = ~resp_id
)
conditional_amces_party
```

```{r}
reference_rows <- conditional_amces_party |>
  separate_wider_delim(contrast, delim = " - ", names = c("level", "ref")) |>
  select(term, level = ref, resp_party) |>
  distinct() |>
  mutate(estimate = 0, conf.low = 0, conf.high = 0)

plot_data <- conditional_amces_party |>
  separate_wider_delim(contrast, delim = " - ", names = c("level", "ref")) |>
  bind_rows(reference_rows) |>
  left_join(attribute_levels_lookup, by = join_by(level))

ggplot(plot_data, aes(x = estimate, y = level_nice, color = resp_party)) +
  geom_vline(xintercept = 0) +
  geom_pointrange(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.15)
  ) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(
    values = party_colors,
    guide = guide_legend(reverse = FALSE)
  ) +
  labs(x = "AMCE", y = "Candidate gender", color = "Respondent party") +
  facet_wrap(vars(term), scales = "free_y", ncol = 2) + 
  theme(legend.position = "bottom")
```

That looks reasonable (and interesting!). For Democratic-identifying respondents, flipping a hypothetical candidate's gender from male to female causes a 7.5 percentage point increase in the probability of being selected. Among Republican-identifying respondents, the effect is negative and causes a 1 percentage point decrease in probability. The effect is statistically significant for Democrats; it's not for Republicans.

It's tempting to look at the gap in AMCEs and say that there's an ≈8.5 percentage point difference between Democrats and Republicans when they're presented with a male vs. female candidate, since that's the distance between the AMCEs (7.5 - (-1)). **But that's wrong!** That distance is an illusion---a mirage caused by the fact that the AMCEs are based on a reference category that is set to zero. Forcing male to 0 for both parties collapses all information about how each party views male candidates.

To get the correct party-based difference in support for candidates, we need to find the party-based marginal means of support for the different candidate genders and then talk about *those* differences:

```{r}
conditional_mms_party <- avg_predictions(
  model_party,
  by = c("gender", "resp_party"),
  vcov = ~resp_id
)
conditional_mms_party
```

```{r}
ggplot(conditional_mms_party, aes(x = estimate, y = gender, color = resp_party)) +
  geom_vline(xintercept = 0.5) +
  geom_pointrange(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.15)
  ) +
  scale_x_continuous(labels = label_percent()) +
  scale_color_manual(
    values = party_colors,
    guide = guide_legend(reverse = FALSE)
  ) +
  labs(x = "Marginal means", y = "Candidate gender", color = "Respondent party") +
  theme(legend.position = "bottom")
```

These marginal means have a direct descriptive interpretation. Republican respondents tend to prefer male candidates (50.5%) to female candidates (49.2%), while Democrats show a stronger preference for female candidates (53.7%) over male candidates (46.3%). These numbers are all on the absolute probability scale with no reference categories---each number tells us the proportion of respondents who would select a candidate with that gender, holding all other attributes constant.

We can measure these cross-party differences directly by calculating the difference in marginal means between respondent parties, separately for each candidate gender:

```{r}
conditional_mms_party_diffs <- avg_comparisons(
  model_party,
  variables = "resp_party",
  by = "gender",
  vcov = ~resp_id
)
conditional_mms_party_diffs
```

These differences are on the probability scale and require no reference category. They answer a straightforward question: for a candidate of a given gender, how much more or less likely are Republicans to select that candidate compared to Democrats? Republicans are ≈4 percentage points more likely than Democrats to select a male candidate and ≈5 percentage points less likely than Democrats to select a female candidate. These are the quantities that the earlier AMCE comparison obscured. When we compared AMCEs across parties, we saw an apparent 8.5 percentage point gap that partially conflated two things: how each party evaluates female candidates *and* how each party evaluates male candidates. By contrast, the marginal mean differences reveal the actual partisan gap: a roughly 4--5 point divergence in each direction from what we'd see if both parties evaluated the candidates identically. Because respondent party is not randomly assigned, these are descriptive differences in preferences rather than causal effects of party identification---but for comparing how groups weight attributes, these marginal mean differences are the correct quantities.

::: {.callout-tip title="Your turn 4"}
The analysis above looked at respondent party identification. Try the same approach using respondent sex (`resp_gender`):

1. Fit a model that interacts `resp_gender` with all candidate attributes (using the same formula structure as `model_party`, but replacing `resp_party` with `resp_gender`).
2. Compute conditional AMCEs for candidate gender by `resp_gender`. What does the apparent difference between female and male respondents look like?
3. Compute conditional marginal means for candidate gender by `resp_gender`. Do female and male respondents evaluate female candidates differently on an absolute scale?
4. Compute the difference in marginal means across `resp_gender` groups for each candidate gender. Are the sex-based respondent differences stronger or weaker than the party-based differences found above?
:::

```{r}
#| label: your-turn-4

# Do stuff here
```

::: {.callout-note title="Your turn 4 answers"}
1. See code below.
2. **The conditional AMCEs look like there's a meaningful gap**: male respondents show only a 2 percentage point preference for female candidates (p = 0.094, not significant), while female respondents show a 6 percentage point preference (p < 0.001). But this apparent gap partly reflects differences in how each group views *male* candidates, not just female candidates.
3. **On the absolute scale**, both groups prefer female candidates, but female respondents do so more strongly: male respondents select male candidates 49.0% and female candidates 51.0% of the time; female respondents select male candidates 47.0% and female candidates 53.1% of the time.
4. **The sex-based differences are weaker than the party-based differences.** Female respondents are ≈2 percentage points less likely than male respondents to select a male candidate and ≈2 percentage points more likely to select a female candidate (both statistically significant). The partisan gaps were roughly twice as large (~4--5 percentage points in each direction).
:::

```{r}
#| label: your-turn-4-answers
#| code-fold: true

model_resp_gender <- lm(
  choice ~ (gender +
    years_politics +
    spouse_occupation +
    occupation +
    n_children +
    age) *
    resp_gender,
  data = candidates
)

# Conditional AMCEs by respondent gender
avg_comparisons(
  model_resp_gender,
  variables = "gender",
  by = "resp_gender",
  vcov = ~resp_id
)

# Conditional marginal means by respondent gender
avg_predictions(
  model_resp_gender,
  by = c("gender", "resp_gender"),
  vcov = ~resp_id
)

# Difference in marginal means across respondent gender groups
avg_comparisons(
  model_resp_gender,
  variables = "resp_gender",
  by = "gender",
  vcov = ~resp_id
)
```

## Oh no! 2: Preference cycles and average feature choice probability (AFCP)

AMCEs are great, but they also implicitly aggregate all sorts of information (that's the "A" part---they average things together). This can be a problem! Respondents saw pairs of candidate profiles, and the context of choices matters and could lead to intransitive or cyclical preferences that we can't detect. For instance, with spouse occupation, it could be that a farmer is preferred over a doctor, a doctor over unmarried, and unmarried over farmer. That's a preference cycle---like rock-paper-scissors!

To fix this, @AbramsonKocakMagazinnik2024 propose a new kind of estimand called the average feature choice probability (AFCP) that accounts for the context of other choices.

To estimate AFCPs, we need to add new columns that show what each respondent task profile was paired against:

```{r}
candidates_alts <- candidates |>
  group_by(resp_id, task) |>
  mutate(across(
    c(gender, years_politics, spouse_occupation, occupation, n_children, age),
    list(alt = \(x) rev(x))
  )) |>
  ungroup()
```

We can see what's happening by looking at the first respondent-task. Among other attributes, the respondent saw a female mayor paired against a female corporate lawyer. The `_alt` versions of the attribute columns here show what the opposing pair was.

```{r}
candidates_alts |>
  filter(resp_id == first(resp_id), task == 1) |>
  select(choice, gender, occupation, gender_alt, occupation_alt)
```

::: {.callout-warning title="More than two profiles"}
We can use `rev()` here because each task contains exactly two profiles, so each profile has a single, well-defined alternative. If the experiment included 3+ profiles per task, this approach would no longer work. There is no unique “other profile” to reverse against, so we must instead expand each task into all pairwise comparisons within the choice set.

For example, with three profiles (A, B, C), each task entails six directed comparisons:

- A vs B  
- A vs C  
- B vs A  
- B vs C  
- C vs A  
- C vs B  

We'd need to expand the data (likely with something like `expand_grid()`) to take into account all the pairwise comparisons within a task. That's left as an exercise to the reader :)
:::

With this new data, we can estimate a model that interacts each attribute with the attribute it was paired with:

```{r}
model_afcp <- lm(
  choice ~ gender * gender_alt +
    years_politics * years_politics_alt +
    spouse_occupation * spouse_occupation_alt +
    occupation * occupation_alt +
    n_children * n_children_alt +
    age * age_alt,
  data = candidates_alts
)
```

To actually calculate the AFCP, we can calculate the predicted probabilities for specific pairings of attribute levels, holding the competing profile fixed and averaging over all the other attributes.

Let's look at the candidate's occupation:

```{r}
avg_predictions(
  model_afcp,
  by = c("occupation", "occupation_alt"),
  newdata = subset(occupation_alt == "Mayor"),  # Competing profile is mayor, for fun
  vcov = ~resp_id
)
```

This is fascinating stuff here! Here's how to interpret these things:

- When mayor is paired against mayor, there's a 50% chance that one of the options will be chosen. That's not surprising---it's the same value, so it has no bearing on respondent choice.
- When teacher is paired against mayor, the teacher option beats mayor 44.3% of the time
- When corporate lawyer is paired against mayor, the lawyer option beats mayor only 36.6% of the time
- Finally, when state legislator is paired against mayor, the state legislator option beats mayor 52.3% of the time

We can also look at differences of each of these pairings against mayors. In direct contests between corporate lawyers and mayors, lawyers are 13.4 percentage points less likely to win:

```{r}
avg_predictions(
  model_afcp,
  by = c("occupation", "occupation_alt"),
  # 0.5 because that's the baseline probability for two choices
  transform = \(x) x - 0.5,
  newdata = subset(occupation_alt == "Mayor"),
  vcov = ~resp_id
)
```

With these differences, we can talk about relative performance too. Lawyers seem to do worse against mayors than teachers do. We can test that formally:

```{r}
#| warning: false

avg_predictions(
  model_afcp,
  by = c("occupation", "occupation_alt"),
  # b2 = lawyer, b1 - teacher
  hypothesis = "b2 - b1 = 0",
  newdata = subset(occupation_alt == "Mayor"),
  vcov = ~resp_id
)
```

This shows that the difference in selection probability between teacher and mayor (0.5 - 0.443, or 0.057) is statistically significantly smaller than the difference between lawyer and mayor (0.5 - 0.366, or 0.134). The gap is -0.0771.

You could spend an entire paper just exploring all the different head-to-head dynamics here!

::: {.callout-tip title="Your turn 5"}
The analysis above explored head-to-head dynamics for candidate occupation. Explore the same for spouse occupation:

1. Compute AFCPs for `spouse_occupation` when the competing profile's spouse occupation is fixed to "Unmarried" using `newdata = subset(spouse_occupation_alt == "Unmarried")`. Which spouse occupation wins most often against an unmarried candidate?
2. Compute the same AFCPs with `transform = \(x) x - 0.5` to center at the 50% baseline. Is the ordering what you'd expect from the simple AMCEs?
3. Repeat the analysis holding the competitor fixed to "Doctor". Does the relative ordering of winner and losers change compared to competing against "Unmarried"?
4. Is there any evidence of a preference cycle? A cycle would require something like: Farmer beats Doctor, Doctor beats Unmarried, but Unmarried beats Farmer.
:::

```{r}
#| label: your-turn-5

# Do stuff here
```

::: {.callout-note title="Your turn 5 answers"}
1. **Farmer and Doctor are essentially tied against Unmarried**, each winning about 57% of direct matchups. This is a mild surprise: the simple AMCEs suggested Farmer (+8 percentage points) had a meaningful edge over Doctor (+6 percentage points) relative to Unmarried, but in actual pairings the two perform nearly identically.
2. **Centered at 0.5, both Farmer (+6.9 percentage points) and Doctor (+7.0 percentage points) beat Unmarried by roughly the same margin.** The AFCP erases the apparent Farmer-vs.-Doctor gap from the simple AMCEs.
3. **The ordering does shift when Doctor is the competitor.** Farmer beats Doctor 53.8% of the time, while Unmarried only wins 43.0% of the time against Doctor. So Farmer still beats Doctor, but the relative advantage of being "married at all" over being unmarried is now clearly visible.
4. **No preference cycle.** All three pairings are visible across the two `avg_comparisons()` chunks: Farmer beats Doctor, Doctor beats Unmarried, and Farmer beats Unmarried---the ordering Farmer > Doctor > Unmarried is transitive. The interesting story is instead how the AFCP reveals that the Farmer-vs.-Doctor gap in the simple AMCEs was overstated: in direct competition, Farmer only edges Doctor by ~4 pp, but both easily beat Unmarried by ~7 pp.
:::

```{r}
#| label: your-turn-5-answers
#| code-fold: true

# AFCPs for spouse occupation competing against "Unmarried"
avg_predictions(
  model_afcp,
  by = c("spouse_occupation", "spouse_occupation_alt"),
  newdata = subset(spouse_occupation_alt == "Unmarried"),
  vcov = ~resp_id
)

# Centered at 0.5
avg_predictions(
  model_afcp,
  by = c("spouse_occupation", "spouse_occupation_alt"),
  transform = \(x) x - 0.5,
  newdata = subset(spouse_occupation_alt == "Unmarried"),
  vcov = ~resp_id
)

# AFCPs for spouse occupation competing against "Doctor"
avg_predictions(
  model_afcp,
  by = c("spouse_occupation", "spouse_occupation_alt"),
  newdata = subset(spouse_occupation_alt == "Doctor"),
  vcov = ~resp_id
)

# Pairwise differences when competing against "Unmarried"
avg_comparisons(
  model_afcp,
  variables = list(spouse_occupation = "pairwise"),
  newdata = subset(spouse_occupation_alt == "Unmarried"),
  vcov = ~resp_id
)

# Pairwise differences when competing against "Doctor"
avg_comparisons(
  model_afcp,
  variables = list(spouse_occupation = "pairwise"),
  newdata = subset(spouse_occupation_alt == "Doctor"),
  vcov = ~resp_id
)
```

## Reflection questions

1. The main lesson from "Oh no! 1" is that comparing conditional AMCEs across subgroups can be misleading. In your own words, why does the AMCE comparison fail? When would comparing AMCEs across groups give you roughly the right answer?

::: {.callout-note title="Answer"}
AMCEs are differences from a reference level---they answer "how much does switching from Male to Female shift the probability of selection?" When you compare this quantity across groups, you are comparing *deviations from a baseline*, but each group's baseline may differ. Democrats' 7.5 percentage point AMCE means that Democrats prefer female candidates 7.5 percentage points over male candidates; Republicans' −1 percentage point AMCE means Republicans are essentially indifferent. The apparent 8.5 percentage point gap across parties conflates two things: how Democrats evaluate female candidates *and* how Democrats evaluate male candidates compared to Republicans. The comparison would give roughly the right answer when both groups have similar absolute support for the reference category---i.e., when the baselines are roughly equal. The marginal mean differences avoid this problem entirely because they express preferences on an absolute probability scale with no hidden reference category.
:::

2. The AFCP estimand accounts for the context of the competing profile. Can you think of a research question where context-sensitivity would matter a lot? Can you think of one where a standard AMCE would be perfectly adequate?

::: {.callout-note title="Answer"}
Context sensitivity matters most when preferences are inherently comparative or non-transitive. For example, voters might consistently prefer "the more moderate candidate" in any given pair---producing cycles where moderate beats conservative, conservative beats liberal, and liberal beats moderate. Studies of racial or gender bias might also find that evaluations of a minority candidate depend heavily on who they are paired against. Standard AMCEs would miss these dynamics entirely.

AMCEs are adequate when preferences are primarily driven by absolute attribute value rather than relative comparison. If you are studying whether more political experience always increases candidate support regardless of what the competitor looks like, the AMCE captures that cleanly. The AMCE is also simpler to interpret and communicate, and is fine for the most common goal of estimating average causal effects of specific attributes.
:::

4. Conjoint designs assume that attribute levels are randomly and independently assigned, including "implausible" combinations like a 29-year-old candidate with 8 years of political experience. Should we worry about this? What would happen to causal inference if respondents mentally discounted implausible profiles?

::: {.callout-note title="Answer"}
This is a real concern, sometimes called the profile realism problem. If respondents mentally disqualify implausible profiles---ignoring stated attributes or refusing to engage seriously with the profile---those observations contribute noise rather than information. AMCEs estimated from the full data may reflect a mixture of "realistic" and "unrealistic" profile responses, potentially distorting estimates.

Randomization still protects the marginal effects of individual attributes from confounding, but the estimates may not represent preferences for realistic profiles. Some researchers address this by constraining the design to only realistic attribute combinations, though this complicates analysis and can introduce other biases. In practice, many conjoint designs include a moderate number of implausible profiles without major problems---but the issue deserves consideration when attributes are strongly correlated in the real world (e.g., age and years of experience genuinely co-vary for real candidates).
:::
